A performance assessment of evolutionary algorithms in volunteer computing environments: the importance of entropy

### 2016-11-15 EvoPar2017/paper.tex
Reviewed abstact and Introduction

\begin{abstract}
In a volunteer distributed computing system, users run a program on
their own machine to contribute to a common effort. If the program is
embedded in a web page, collaboration is straightforward, but also ephemeral. In
this paper, we analyze a volunteer evolutionary computing system called
NodIO, by running several experiments, some of them massive. Our
objective is to discover rules that encourage volunteer
participation and also the interplay of these contributions with the
dynamics of the algorithm itself, making it more or less efficient. 
We will show different measures of
participation and contribution to the algorithm and how different 
volunteer usage patterns and tweaks in the algorithm, such as
restarting clients when a solution has been found, contribute to 
improvements and leveraging of these contributions. We will also try
to find out what is the key factor in the early termination of the
experiments, measuring entropy in the contributions and other large
scale indicators. 
\end{abstract}

@@

Using this approach for creating distributed experiments is called {\em
  volunteer}, {\em cycle-scavenging} or {\em opportunistic} computing
\cite{sarmenta2001volunteer} and it dates back, in different shapes
and underlying mechanisms, to the origin of the web \cite{david-seti:home}. Our interest
here, however, is to use it as a resource for evolutionary
computation, as our group has done for a long time \cite{jj-ppsn98}.

In this line of research that uses volunteer computing for
evolutionary algorithms, there are several pending issues. The first
and maybe most important is approaching volunteer computing as a
socio-technical system \cite{vespignani2009predicting} which integrates
user decisions and behavioral patterns in the system model; this
includes trying to optimize the number of users in a particular
experiment. The second line of research, although related to the
first, is more focused on the evolutionary algorithm itself and how
different design decisions will affect its performance.
% Is it ok that we say that the first line of research is the most important... but then we focus on the second?
We have approached the first issue in our previous work \cite{jj-ppsn98anonymous},
% I have justified to focus, in this case, on the second
but in this paper our focus will be in the second aspect: we will try to
design a decentralized system that, at the same time, is able to use
all available resources for finding the solution of an evolutionary algorithm. This design will be done incrementally by
changing client and the server and measuring its
impact on the overall performance: time and evaluations needed to find
the solution. Eventually, we want to find a system that, whatever the
number of users available to perform the experiment, is able to
maximize their contribution to the evolutionary algorithm, at the same
time that the evolutionary algorithm itself makes the most of those
contributions and is able to find
the solution to the problem in a minimum time and with the least
number of contributions.

>>>>

\begin{abstract}
In a volunteer distributed computing system, users run a program on
their own machine to contribute to a common effort. If the program is
embedded in a web page, collaboration is straightforward, but also ephemeral. In
this paper, we analyze a volunteer evolutionary computing system called
NodIO, by running several experiments, some of them massive. Our
objective is to discover rules that encourage volunteer
participation and also the interplay of these contributions with the
dynamics of the algorithm itself, making it more or less efficient. % efficient enough? Somehow saying "more or less efficient" does not seem "our work is awesome" to me.
We will show different measures of
participation and contribution to the algorithm, as well as how different 
volunteer usage patterns and tweaks in the algorithm, such as
restarting clients when a solution has been found, contribute to 
improvements and leveraging of these contributions. We will also try
to find out what is the key factor in the early termination of the
experiments, measuring entropy in the contributions and other large
scale indicators.
\end{abstract}

@@

Using this approach for creating distributed experiments is called {\em
  volunteer}, {\em cycle-scavenging}, or {\em opportunistic} computing
\cite{sarmenta2001volunteer} and it dates back, in different shapes
and underlying mechanisms, to the origin of the web \cite{david-seti:home}. Our interest
here, however, is to use it as a resource for evolutionary
computation, as our group has done for a long time \cite{jj-ppsn98}. % Not anonymised

In this line of research that uses volunteer computing for
evolutionary algorithms, there are several pending issues. The first
and maybe most important is approaching volunteer computing as a
socio-technical system \cite{vespignani2009predicting} which integrates
user decisions and behavioral patterns in the system model; this
includes trying to optimize the number of users in a particular
experiment. The second line of research, although related to the
first, is more focused on the evolutionary algorithm itself and how
different design decisions will affect its performance.
% Is it ok that we say that the first line of research is the most important... but then we focus on the second?
We have approached the first issue in our previous work \cite{jj-ppsn98anonymous},
% I have justified to focus, in this case, on the second
but in this paper our focus will be in the second aspect: we will try to
design a decentralized system that, at the same time, is able to use
all available resources for finding the solution of an evolutionary algorithm. This design will be done incrementally by
changing client and the server and measuring its
impact on the overall performance: time and evaluations needed to find
the solution. Eventually, we want to find a system that, whatever the
number of users available to perform the experiment, is able to
maximize their contribution to the evolutionary algorithm, at the same
time that the evolutionary algorithm itself makes the most of those
contributions and is able to find
the solution to the problem in a minimum time, with the least
number of contributions.
